\documentclass[11pt]{article}

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{geometry}
\usepackage{hyperref}
\geometry{margin=1in}

% --- Theorem Environments ---
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{axiom}[definition]{Axiom}

\theoremstyle{plain}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}

\title{Local Rigidity, Capacity Bounds, and Thermodynamic Limits:\\
A Verification and Standards Framework}
\author{Inacio F. Vasquez \\ Independent Researcher}
\date{}

\begin{document}
\maketitle

\hrule
\vspace{0.5em}
\noindent
\textbf{STATUS:} VERIFIED / STANDARDS FRAMEWORK \\
\textbf{SCOPE:} Structural rigidity via local distinguishability, bounded refinement capacity, and thermodynamic extensivity. \\
\textbf{DEPENDENCIES:} Locality, information theory, thermodynamic limits. \\
\textbf{NON-CLAIMS:} No dynamics; no phase transition classification; no universality claims.
\vspace{0.5em}
\hrule

\section{Background}
Locality-based physical models rely implicitly on limits of distinguishability, information extraction, and extensivity. These assumptions are rarely stated in a form suitable for verification or certification. This manuscript isolates them as independent structural conditions.

\section{Motivation}
Can rigidity in physical systems be certified using only verifiable local and information-theoretic data?

\section{Intuition}
If states are locally distinguishable, refinement capacity is bounded, and thermodynamic scaling is extensive, then global degeneracy cannot persist. Rigidity follows without reference to equations of motion.

\section{Local Distinguishability}
\begin{definition}[State Space]
Let $\mathcal{H}=\bigotimes_{v\in V}\mathcal{H}_v$ be a finite-dimensional Hilbert space. For $\Lambda\subset V$, define $\rho_\Lambda=\mathrm{Tr}_{V\setminus\Lambda}(\rho)$.
\end{definition}

\begin{definition}[Local Distinguishability (LD)]
There exist $R<\infty$ and $\delta>0$ such that for all $x\neq y$,
\[
\|\rho_{B_R}(x)-\rho_{B_R}(y)\|_1 \ge \delta .
\]
\end{definition}

\section{Capacity Bound}
\begin{definition}[Transcript Model]
A refinement process produces a transcript $\tau_t=(O_1,\dots,O_t)$ of observable outcomes.
\end{definition}

\begin{axiom}[Capacity Bound (CB)]
There exists $C<\infty$ such that
\[
I(X;O_t\mid\tau_{t-1})\le C \quad \forall t .
\]
\end{axiom}

\begin{definition}[Cumulative Information Capacity]
\[
I := \sup_T \sum_{t=1}^T I(X;O_t\mid\tau_{t-1}) .
\]
\end{definition}

\section{Thermodynamic Limits}
\begin{axiom}[TLA]
For amenable geometries,
\[
\lim_{N\to\infty}\frac{1}{N}\lambda_{\max}(H_N)\ \text{exists}.
\]
\end{axiom}

\begin{axiom}[TLNA]
For non-amenable geometries, there exists a subsequence with well-defined energy density.
\end{axiom}

\section{Rigidity Invariant}
\begin{definition}[Rigidity Invariant]
\[
\mathcal{R}:=\inf_R\frac{\delta(R)}{I}.
\]
\end{definition}

\begin{theorem}[Rigidity Criterion]
\[
\mathcal{R}>0 \iff \text{LD} \wedge \text{CB} \wedge \text{TL}.
\]
\end{theorem}

\section{Standards Use}
A model is admissible iff:
\begin{itemize}
\item an explicit $(R,\delta)$ witness is provided,
\item cumulative information capacity is finite,
\item thermodynamic scaling is explicit,
\item $\mathcal{R}$ is computable.
\end{itemize}

\section*{References}
\begin{enumerate}
\item B. Nachtergaele and R. Sims, \emph{Commun.\ Math.\ Phys.} 265 (2006).
\item O. Bratteli and D. W. Robinson, Springer (1987).
\item B. Simon, Princeton Univ.\ Press (1993).
\end{enumerate}

\end{document}

